---
title: "Analyze Zebra Box"
---
## Usage: This script analyzes the output of zebrafish video recording data taken by ZebraBox and displays summary data such as movement over time, average distance moved, average duration moved, and average speed by experimental condition group. 

## REQUIRED: Prior to running this code, please generate an .xlsx/.xls file with a column labeled "Type" including a label for the injected/experimental condition of each fish this is the 'wells' file found in the 'Load in data' section

```{r}
# Install Necessary Packages -------------------------------------------------
install.packages("xlsx", dependencies = TRUE)
install.packages("ggplot2", dependencies = TRUE)
install.packages("ggpubr", dependencies = TRUE)
install.packages("ggsignif")
install.packages("dplyr")
install.packages("readxl")
```

```{r}
# Load Necessary Packages -------------------------------------------------
library("readxl")
library("xlsx")
library("ggplot2")
library("ggpubr")
library("ggsignif")
library("dplyr")
```

```{r}
# Customize Gene Name Here ------------------------------------------------
gene <- "Example Data" #Change this line to reflect your gene name

comp = list(c("knockout", "wildtype"), c("knockout", "injection_control"), c("injection_control", "wildtype")) #Designate which fish experimental condition group you would like to compare for p-value calculation

# Customize Save File -----------------------------------------------------
gene_name <- gsub(" ", "_", gene, fixed = TRUE)
save_name <- paste("ZBox_",gene_name,".pdf",sep="") #Change file name here - default save will be "FishInspector_gene.pdf"
```

```{r}
# Load in data: Customize Files Locations Here ----------------------------
file_1 <- read_excel("~/Desktop/Analyze_Zebra_Box/example_data_1.xlsx") #Change this line to your own file path
wells_file_1 <- read_excel("~/Desktop/Analyze_Zebra_Box/example_data_wells_1.xls") #Please provide an .xlsx/.xls file with well designations and fish experimental condition group (ex: A1 wildtype)

Type <-rep((unlist(wells_file_1[,2])),times=240)
file_1_type <- cbind(Type, file_1)

all_type <- file_1_type 

all_type <- all_type[all_type$Type != "EMPTY",] #Removes all wells labeled as 'EMPTY'
```

```{r}
#Calculate Average Distance and Duration for File 1 ----------------------
file_1_df <- file_1 %>%
  select(location, lardist, lardur) %>%
  group_by(location) %>%
  summarize(Average_Distance=mean(lardist), Average_Duration=mean(lardur))

file_1_df <- cbind(wells_file_1, file_1_df)

all_files <- file_1_df 

#Remove Spurious Data
all_files <- all_files[all_files$Type != "EMPTY",] # Remove EMPTY wells
#all_files <- all_files[all_files$Average_Duration >50,] # Remove likely dead fish

#Group Fish Longitudinally by Experimental Condition (Type)
options(dplyr.summarise.inform = FALSE)
group_type <- all_files %>%
  group_by(Type) %>%
  summarize(Average_Distance=mean(Average_Distance), Err_Dist=sd(Average_Distance)/sqrt(n()), Average_Duration=mean(Average_Duration), Err_Dur=sd(Average_Duration)/sqrt(n()), Distance_Duration=mean(Average_Distance/Average_Duration), Err_Dist_Durr=sd(Average_Distance/Average_Duration)/sqrt(n()), n = n()) 

group_type <- all_type %>%
  group_by(Type, start) %>%
  summarize(Average_Distance=mean(lardist), n=n())

fish_numbs <- data.frame(c(unique(group_type[,1]),unique(group_type[,4])))
fish_numbs <- paste(unlist(split(fish_numbs, seq(nrow(fish_numbs)))), collapse = " ")
```

```{r}
#If needed, Load in a second Zebra Box file ---------------------------
file_2 <- read_excel("~/Desktop/Analyze_Zebra_Box/example_data_2.xlsx")
wells_file_2 <- read_excel("~/Desktop/Analyze_Zebra_Box/example_data_wells_2.xls")

Type <-rep((unlist(wells_file_2[,2])),times=240)
file_2_type <- cbind(Type, file_2)

all_type <- rbind(file_1_type,file_2_type) 

all_type <- all_type[all_type$Type != "EMPTY",] #Removes all wells labeled as 'EMPTY'
```

```{r}
#Calculate Average Distance and Duration for File 2 ----------------------
file_2_df <- file_2 %>%
  select(location, lardist, lardur) %>%
  group_by(location) %>%
  summarize(Average_Distance=mean(lardist), Average_Duration=mean(lardur))

file_2_df <- cbind(wells_file_2, file_2_df)

#Merge Files + Add additional files as needed ---------------------------
all_files <- rbind(file_1_df,file_2_df) 

#Remove Spurious Data
all_files <- all_files[all_files$Type != "EMPTY",] # Remove EMPTY wells
#all_files <- all_files[all_files$Average_Duration >50,] # Remove likely dead fish

#Group Fish Longitudinally by Experimental Condition (Type)
group_type <- all_files %>%
  group_by(Type) %>%
  summarize(Average_Distance=mean(Average_Distance), Err_Dist=sd(Average_Distance)/sqrt(n()), Average_Duration=mean(Average_Duration), Err_Dur=sd(Average_Duration)/sqrt(n()), Distance_Duration=mean(Average_Distance/Average_Duration), Err_Dist_Durr=sd(Average_Distance/Average_Duration)/sqrt(n()), n = n()) 

group_type <- all_type %>%
  group_by(Type, start) %>%
  summarize(Average_Distance=mean(lardist), n=n())

fish_numbs <- data.frame(c(unique(group_type[,1]),unique(group_type[,4])))
fish_numbs <- paste(unlist(split(fish_numbs, seq(nrow(fish_numbs)))), collapse = " ")
```

```{r dev = "pdf", dpi = 300}
# Generate Box Plots ------------------------------------------------------
# Average  Distance Moved
plot1 <- ggplot(all_files, mapping=aes(x=Type, y=Average_Distance)) +
  geom_boxplot() + theme_bw() + 
  ggtitle("Average Distance Moved") + theme(plot.title = element_text(hjust = 0.5)) + ylab("Distance") + xlab("") + 
  geom_signif(comparisons = comp, map_signif_level = FALSE, step_increase = 0.1)

#Average Duration Moved
plot2 <- ggplot(all_files, mapping=aes(x=Type, y=Average_Duration)) +
  geom_boxplot() + theme_bw() + 
  ggtitle("Average Duration Moved") + theme(plot.title = element_text(hjust = 0.5)) + ylab("Duration") + xlab("") + 
  geom_signif(comparisons = comp, map_signif_level = FALSE, step_increase = 0.1)

#Average Distance/Duration Moved
plot3 <- ggplot(all_files, mapping=aes(x=Type, y=Average_Distance/Average_Duration )) +
  geom_boxplot() + theme_bw() + 
  ggtitle("Average Speed") + theme(plot.title = element_text(hjust = 0.5)) + ylab("Distance/Duration") + xlab("") + 
  geom_signif(comparisons = comp, map_signif_level = FALSE, step_increase = 0.1)


#Distance Over Time
plot4 <- ggplot(group_type, mapping=aes(x=start, y=Average_Distance, color=Type))+
geom_smooth() + theme_bw() + 
geom_vline(xintercept = 3600, linetype="dashed", color = "grey") +
ggtitle("Distance Over Time") + theme(plot.title = element_text(hjust = 0.5)) + ylab("Distance Moved") + xlab("Time (seconds)")
  

#Display Plots
fig <- ggarrange(plot4,                                     
        ggarrange(plot1, plot2, plot3, ncol = 3, labels = c("B","C","D")),
        nrow = 2, labels = "A" )
plot(fig)
annotate_figure(fig, top = gene, bottom = fish_numbs)

ggsave(save_name, plot = last_plot(), width = 12, height = 12, dpi = 300) #Change dimensions of saved PDF here

dev.off()          
```
